---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

descriptr: Generate descriptive statistics and explore distributions 
----------

**Author:** [Aravind Hebbali](http://www.aravindhebbali.com)<br/>
**License:** [MIT](https://opensource.org/licenses/MIT)

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/descriptr)](https://cran.r-project.org/package=descriptr)
[![Travis-CI Build Status](https://travis-ci.org/rsquaredacademy/descriptr.svg?branch=master)](https://travis-ci.org/rsquaredacademy/descriptr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rsquaredacademy/descriptr?branch=master&svg=true)](https://ci.appveyor.com/project/rsquaredacademy/descriptr) 
[![](https://cranlogs.r-pkg.org/badges/grand-total/descriptr)](https://cran.r-project.org/package=descriptr) [![Coverage status](https://codecov.io/gh/rsquaredacademy/descriptr/branch/master/graph/badge.svg)](https://codecov.io/github/rsquaredacademy/descriptr?branch=master) ![](https://img.shields.io/badge/lifecycle-maturing-blue.svg)



## Overview

The goal of descriptr is to ease the process of generating descriptive statistics and exploring statistical distributions.

## Installation

```{r cran-installation, eval = FALSE}
# install descriptr from CRAN
install.packages("descriptr")

# the development version from github
# install.packages("devtools")
devtools::install_github("rsquaredacademy/descriptr")
```

## Features

### Descriptive Statistics

- Summary Statistics
- Two Way Tables
- One Way Table
- One Way Table (Continuous Data)
- Group Wise Summary
- Multiple Column Descriptive Statistics
- Multiple One Way Tables
- Multiple Two Way Tables


### Explore Distributions

- Normal
- Binomial
- Chi Square
- F
- t


## Shiny App

Use `ds_launch_shiny_app()` to explore the package using a shiny app.

## Vignettes

- [Descriptive Statistics](http://www.rsquaredacademy.com/descriptr/articles/descriptive-stats.html)
- [Statistical Distributions](http://www.rsquaredacademy.com/descriptr/articles/distributions.html)

## Usage

We will use a modified version of the `mtcars` data set in the below examples. The only difference between the data sets is related to the 
variable types.

```{r load, eval=TRUE, echo=FALSE}
library(descriptr)
```

```{r egdata}
str(mtcarz)
```

##### Summary Statistics

```{r descript1}
ds_summary_stats(mtcarz, mpg)
```

##### Two Way Table

```{r descript2}
ds_cross_table(mtcarz, cyl, gear)
```

##### One Way Table

```{r descript3}
ds_freq_table(mtcarz, cyl)
```

##### One Way Table (Continuous Data)

```{r descript4}
ds_freq_cont(mtcarz, mpg)
```

##### Group Summary

```{r descript5}
ds_group_summary(mtcarz, cyl, mpg)
```

##### Multiple Variable Statistics

```{r descriptr6}
ds_multi_stats(mtcarz, mpg, disp, hp)
```


Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project you agree to abide by its terms.
