---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# descriptr 

> Generate descriptive statistics

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/descriptr)](https://cran.r-project.org/package=descriptr) [![cran checks](https://cranchecks.info/badges/summary/descriptr)](https://cran.r-project.org/web/checks/check_results_descriptr.html)
[![Travis-CI Build Status](https://travis-ci.org/rsquaredacademy/descriptr.svg?branch=master)](https://travis-ci.org/rsquaredacademy/descriptr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rsquaredacademy/descriptr?branch=master&svg=true)](https://ci.appveyor.com/project/rsquaredacademy/descriptr) 
[![](https://cranlogs.r-pkg.org/badges/grand-total/descriptr)](https://cran.r-project.org/package=descriptr) [![Coverage status](https://codecov.io/gh/rsquaredacademy/descriptr/branch/master/graph/badge.svg)](https://codecov.io/github/rsquaredacademy/descriptr?branch=master) ![](https://img.shields.io/badge/lifecycle-maturing-blue.svg)


## Installation

```{r cran-installation, eval = FALSE}
# Install release version from CRAN
install.packages("descriptr")

# Install development version from GitHub
# install.packages("devtools")
devtools::install_github("rsquaredacademy/descriptr")
```

## Articles

- [Descriptive Statistics](https://descriptr.rsquaredacademy.com/articles/descriptive-stats.html)
- [Statistical Distributions](https://descriptr.rsquaredacademy.com/articles/distributions.html)

## Usage

We will use a modified version of the `mtcars` data set in the below examples. The only difference between the data sets is related to the 
variable types.

```{r load, eval=TRUE, echo=FALSE}
library(descriptr)
```

```{r egdata}
str(mtcarz)
```

#### Summary Statistics

```{r descript1}
ds_summary_stats(mtcarz, mpg)
```

#### Two Way Table

```{r descript2}
ds_cross_table(mtcarz, cyl, gear)
```

#### One Way Table

```{r descript3}
ds_freq_table(mtcarz, cyl)
```

#### One Way Table (Continuous Data)

```{r descript4}
ds_freq_cont(mtcarz, mpg)
```

#### Group Summary

```{r descript5}
ds_group_summary(mtcarz, cyl, mpg)
```

#### Multiple Variable Statistics

```{r descriptr6}
ds_multi_stats(mtcarz, mpg, disp, hp)
```

## Features

- Summary statistics
- Two way tables
- One way table
- One way table (Continuous Data)
- Group wise summary
- Multiple variable statistics
- Multiple one way tables
- Multiple two way tables

## Getting Help

If you encounter a bug, please file a minimal reproducible example using 
[reprex](https://reprex.tidyverse.org/index.html) on github. For questions and 
clarifications, use [StackOverflow](https://stackoverflow.com/).

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project you agree to abide by its terms.
