---
title: "Introduction to descriptr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to descriptr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE}
library(descriptr)
library(dplyr)
```

## Introduction

Descriptive statistics are used to summarize data. It enables us to present the
data in a more meaningful way and to discern any patterns existing in the data. 
They can be useful for two purposes:

- provide basic information about variables in a data set
- highlight potential relationships between variables

This document introduces you to a basic set of functions that describe data. 
There is a second vignette which provides details about functions which help 
visualize statistical distributions.


## Data Screening

The `screener` function will screen a data set and return the following:
- Column/Variable Names
- Data Type
- Levels (in case of categorical data)
- Number of missing observations
- % of missing observations


```{r screener}
mt <- mtcars
mt[, c(2, 8:11)] <- lapply(mt[, c(2, 8:11)], factor)
mt[sample(1:nrow(mt), 12), sample(1:ncol(mt), 6)] <- NA
screener(mt)
```


## Summary Statistics

The `ds_summary_stats` function returns a comprehensive set of statistics for 
**continuous** data. 

```{r summary}
ds_summary_stats(mtcars$mpg)
```


## Cross Tabulation

The `ds_cross_table` function creates two way tables of categorical variables. It
is **not necessary** to coerce a variable to type `factor`.

```{r cross}
ds_cross_table(mtcars$cyl, mtcars$gear)
```

A plot method has been defined which will generate:


### Grouped Bar Plots

```{r cross_group, fig.width=7, fig.height=7, fig.align='centre'}
k <- ds_cross_table(mtcars$cyl, mtcars$gear)
plot(k, beside = TRUE)
```


### Stacked Bar Plots

```{r cross_stack, fig.width=7, fig.height=7, fig.align='centre'}
k <- ds_cross_table(mtcars$cyl, mtcars$gear)
plot(k)
```


### Proportional Bar Plots

```{r cross_prop, fig.width=7, fig.height=7, fig.align='centre'}
k <- ds_cross_table(mtcars$cyl, mtcars$gear)
plot(k, proportional = TRUE)
```


### Mosaic Plots

Mosaic plots can be created using the `mosaciplot` method.

```{r mosaic, fig.width=7, fig.height=7, fig.align='centre'}
k <- ds_cross_table(mtcars$cyl, mtcars$gear)
mosaicplot(k)
```


## Frequency Table (Categorical Data)

The `ds_freq_table()` function creates frequency tables for categorical variables. 

```{r ftable}
mt <- mtcars
mt$cyl <- as.factor(mt$cyl)
ds_freq_table(mt$cyl)
```


### Bar Plot

A barplot method has been defined.

```{r ftable_bar, fig.width=7, fig.height=7, fig.align='centre'}
mt <- mtcars
mt$cyl <- as.factor(mt$cyl)
k <- ds_freq_table(mt$cyl)
barplot(k)
```


## Frequency Table (Continuous Data)

The `ds_freq_cont` function creates frequency tables for continuous variables. The
default number of intervals is 5.

```{r fcont}
ds_freq_cont(mtcars$mpg, 4)
```


### Histogram

A `hist` method has been defined.

```{r fcont_hist, fig.width=7, fig.height=7, fig.align='centre'}
k <- ds_freq_cont(mtcars$mpg, 4)
hist(k)
```


## Group Summary

The `ds_group_summary()` function returns descriptive statistics of a continuous
variable for the different levels of a categorical variable.

```{r gsummary}
mt <- mtcars
mt$cyl <- as.factor(mt$cyl)
ds_group_summary(mt$cyl, mt$mpg)
```


### Box Plot

A `boxplot()` method has been defined.

```{r gsum_boxplot, fig.width=7, fig.height=7, fig.align='centre'}
mt <- mtcars
mt$cyl <- as.factor(mt$cyl)
k <- ds_group_summary(mt$cyl, mt$mpg)
boxplot(k)
```

## Multi Column Statistics

Use `multistats()` to generate descriptive/summary statistics for multiple columns of a tibble/data.frame.

```{r multistats}
multistats(mtcars, mpg, disp, hp, wt, qsec)
```

## Multiple One Way Tables

The `ds_oway_tables()` function creates multiple one way tables by creating a 
frequency table for each categorical variable in a data frame.

```{r oway}
mt <- mtcars
mt[, c(2, 8:11)] <- lapply(mt[, c(2, 8:11)], factor)
ds_oway_tables(mt)
```


## Multiple Two Way Tables

The `ds_tway_tables()` function creates multiple two way tables by creating a cross
table for each unique pair of categorical variables in a data frame.

```{r tway}
mt <- mtcars
mt[, c(2, 8:10)] <- lapply(mt[, c(2, 8:10)], factor)
ds_tway_tables(mt)
```

## Multiple Variable Statistics 

The `ds_multi_stats()` function generates summary/descriptive statistics for multiple variables in a data frame/tibble.

```{r multistats}
ds_multi_stats(mtcars, mpg, disp, hp)
```



