---
title: "Introduction to descriptr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to descriptr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE}
library(descriptr)
library(dplyr)
```

Descriptive statistics are used to summarize data. It enables us to present the
data in a more meaningful way and to discern any patterns existing in the data. 
They can be useful for two purposes:

- provide basic information about variables in a data set
- highlight potential relationships between variables

This document introduces you to a basic set of functions which describe the data 
in a comprehensive manner. There is a second vignette which provides details 
about function which help visualize statistical distributions.

## Overview

In this section, we will look at functions available in the `descriptr` package
to get an overview of a data set. Below are the functions available:

- `screener`
- `group_summary`

```{r screener}
mt <- mtcars
mt[, c(2, 8:11)] <- lapply(mt[, c(2, 8:11)], factor)
mt[sample(1:nrow(mt), 12), c(2, 4, 7:10)] <- NA
screener(mt)
```

```{r screener_plot, fig.width=5, fig.height=5, fig.align='center'}
mt <- mtcars
mt[, c(2, 8:11)] <- lapply(mt[, c(2, 8:11)], factor)
mt[sample(1:nrow(mt), 12), c(2, 4, 7:10)] <- NA
plot(screener(mt))
```

## Univariate Analysis

In this section, we will look at functions available in the `descriptr` package
to describe a single variable. Below are the functions available:

- `summary_stats`
- `freq_cont`
- `freq_table`
- `oway_tables`
- `skewness`
- `kurtosis`
- `stat_cvar`
- `stat_mode`
- `stat_range`
- `stat_mdev`
- `stat_css`
- `gmean`
- `hmean`

## Bivariate Analysis

In this section, we will look at functions available in the `descriptr` package
to describe the relationship between variables. Below are the functions available:

- `cross_table`
- `tway_tables`
