---
title: "Exploring Distributions"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE}
library(descriptr)
```

In exploring statistical distributions, we focus on the following:

- what influences the shape of a distribution
- calculate probability from a given quantile
- calculate quantiles out of given probability

To explore the above 3 concepts, we have defined functions for the following
distributions:

- Normal
- Binomial
- Chi Square
- F
- t

## Normal Distribution

### Distribution Shape

Visualize how changes in mean and standard deviation affect the shape of the 
normal distribution.

##### Input
- mean: mean of the normal distribution
- sd: standard deviation of the normal distribution

##### Output
- Normal distribution plot

```{r norm_plot, fig.width=7, fig.height=7, fig.align='centre'}
norm_plot(mean = 2, sd = 0.6)
```

### Percentiles

#### Calculate and visualize quantiles out of given probability.

##### Input
- probs: a probability value
- mean: mean of the normal distribution
- sd: standard deviation of the normal distribution
- type: lower/upper tail

Suppose X, the grade on a exam, is normally distributed with mean 60 
and standard deviation 3. The teacher wants to give 10% of the class an A. 
What should be the cutoff to determine who gets an A?

```{r norm_per1, fig.width=7, fig.height=7, fig.align='centre'}
norm_per(0.10, 60, 3, 'upper')
```

The teacher wants to give lower 15% of the class a D. What cutoff should the 
teacher use to determine who gets an D?

```{r norm_per2, fig.width=7, fig.height=7, fig.align='centre'}
norm_per(0.85, 60, 3, 'lower')
```

The teacher wants to give middle 50% of the class a B. What cutoff should the 
teacher use to determine who gets an B?

```{r norm_per3, fig.width=7, fig.height=7, fig.align='centre'}
norm_per(0.5, 60, 3, 'both')
```

###  Probabilities

#### Calculate and visualize probability from a given quantile

##### Input
- perc: a quantile value
- mean: mean of the normal distribution
- sd: standard deviation of the normal distribution
- type: lower/upper/both tail

Let X be the IQ of a randomly selected student of a school. Assume X ~ N(90, 4). 
What is the probability that a randomly selected student has an IQ below 80?

```{r norm_prob1, fig.width=7, fig.height=7, fig.align='centre'}
norm_prob(80, mean = 90, sd = 4)
```

What is the probability that a randomly selected student has an IQ above 100?

```{r norm_prob2, fig.width=7, fig.height=7, fig.align='centre'}
norm_prob(100, mean = 90, sd = 4, type = 'upper')
```

What is the probability that a randomly selected student has an IQ 
between 85 and 100?

```{r norm_prob3, fig.width=7, fig.height=7, fig.align='centre'}
norm_prob(c(85, 100), mean = 90, sd = 4, type = 'both')
```

## Binomial Distribution

### Distribution Shape

Visualize how changes in number of trials and the probability of success affect
the shape of the binomial distribution.

```{r binom_plot, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_plot(10, 0.3)
```

### Percentiles 

#### Calculate and visualize quantiles out of given probability

##### Input
- p: a single aggregated probability of multiple trials
- n: the number of trials
- tp: the probability of success in a trial
- type: lower/upper tail


```{r binom_per1, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_perc(10, 0.5, 0.05)
```

```{r binom_per2, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_perc(10, 0.5, 0.05, 'upper')
```

### Probabilities

#### Calculate and visualize probability from a given quantile

##### Input
- p: probability of success
- n: the number of trials
- s: number of success in a trial
- type: lower/upper/interval/exact tail


Assume twenty-percent (20%) of Magemill have no health insurance. Randomly 
sample n = 12 Magemillians. Let X denote the number in the sample with no 
health insurance. What is the probability that exactly 4 of the 15 sampled 
have no health insurance?

```{r binom_prob1, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_prob(12, 0.2, 4, type = 'exact')
```

What is the probability that at most one of those sampled has no health 
insurance?

```{r binom_prob2, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_prob(12, 0.2, 1, 'lower')
```

What is the probability that more than seven have no health insurance?

```{r binom_prob3, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_prob(12, 0.2, 8, 'upper')
```

What is the probability that fewer than 5 have no health insurance?

```{r binom_prob4, fig.width=7, fig.height=7, fig.align='centre'}
dist_binom_prob(12, 0.2, c(0, 4), 'interval')
```


## Chi Square Distribution

### Distribution Shape

Visualize how changes in degrees of freedom affect the shape of the chi square 
distribution.

```{r chi_plot, fig.width=7, fig.height=7, fig.align='centre'}
dist_chi_plot(df = 5)
dist_chi_plot(df = 5, normal = TRUE)
```

### Percentiles

#### Calculate quantiles out of given probability

##### Input
- probs: a probability value
- df: degrees of freedom
- type: lower/upper tail

Let X be a chi-square random variable with 8 degrees of freedom. What is the 
upper fifth percentile? 

```{r chi_per1, fig.width=7, fig.height=7, fig.align='centre'}
dist_chi_perc(0.05, 8, 'upper')
```

What is the tenth percentile?

```{r chi_per2, fig.width=7, fig.height=7, fig.align='centre'}
dist_chi_perc(0.10, 8, 'lower')
```

### Probability

#### Calculate probability from a given quantile.

##### Input
- perc: a quantile value
- df: degrees of freedom
- type: lower/upper tail


What is the probability that a chi-square random variable with 12 degrees of 
freedom is greater than 8.79?

```{r chi_prob1, fig.width=7, fig.height=7, fig.align='centre'}
dist_chi_prob(8.79, 12, 'upper')
```

What is the probability that a chi-square random variable with 12 degrees of 
freedom is greater than 8.62?

```{r chi_prob2, fig.width=7, fig.height=7, fig.align='centre'}
dist_chi_prob(8.62, 12, 'lower')
```

## F Distribution

### Distribution Shape

Visualize how changes in degrees of freedom affect the shape of the F 
distribution.

```{r f_plot, fig.width=7, fig.height=7, fig.align='centre'}
f_plot()
f_plot(6, 10, normal = TRUE)
```

### Percentiles

#### Calculate quantiles out of given probability

##### Input
- probs: a probability value
- num_df: nmerator degrees of freedom
- den_df: denominator degrees of freedom
- type: lower/upper tail


Let X be an F random variable with 4 numerator degrees of freedom and 5 
denominator degrees of freedom. What is the upper twenth percentile?

```{r f_per1, fig.width=7, fig.height=7, fig.align='centre'}
f_per(0.20, 4, 5, 'upper')
```

What is the 35th percentile?

```{r f_per2, fig.width=7, fig.height=7, fig.align='centre'}
f_per(0.35, 4, 5, 'lower')
```

### Probabilities

#### Calculate probability from a given quantile.

##### Input
- perc: a quantile value
- num_df: nmerator degrees of freedom
- den_df: denominator degrees of freedom
- type: lower/upper tail


What is the probability that an F random variable with 4 numerator degrees of 
freedom and 5 denominator degrees of freedom is greater than 3.89?

```{r f_prob1, fig.width=7, fig.height=7, fig.align='centre'}
f_prob(3.89, 4, 5, 'upper')
```

What is the probability that an F random variable with 4 numerator degrees of 
freedom and 5 denominator degrees of freedom is less than 2.63?

```{r f_prob2, fig.width=7, fig.height=7, fig.align='centre'}
f_prob(2.63, 4, 5, 'lower')
```


## t Distribution

### Distribution Shape

Visualize how degrees of freedom affect the shape of t distribution.

```{r t_plot, fig.width=7, fig.height=7, fig.align='centre'}
t_plot(df = 8)
```

### Percentiles

#### Calculate quantiles out of given probability

##### Input
- probs: a probability value
- df: degrees of freedom
- type: lower/upper/both tail


What is the upper fifteenth percentile? 

```{r t_per1, fig.width=7, fig.height=7, fig.align='centre'}
t_per(0.15, 8, 'upper')
```

What is the eleventh percentile? 

```{r t_per2, fig.width=7, fig.height=7, fig.align='centre'}
t_per(0.11, 8, 'lower')
```

What is the area of the curve that has 95% of the t values? 

```{r t_per3, fig.width=7, fig.height=7, fig.align='centre'}
t_per(0.8, 8, 'both')
```


### Probabilities

#### Calculate probability from a given quantile.

##### Input
- perc: a quantile value
- df: degrees of freedom
- type: lower/upper/interval/both tail


Let T follow a t-distribution with r = 6 df.  

What is the probability that the value of T is less than 2?
 
```{r t_prob1, fig.width=7, fig.height=7, fig.align='centre'}
t_prob(2, 6, 'lower')
```

What is the probability that the value of T is greater than 2?
 
```{r t_prob2, fig.width=7, fig.height=7, fig.align='centre'}
t_prob(2, 6, 'upper')
```

What is the probability that the value of T is between -2 and 2?
 
```{r t_prob3, fig.width=7, fig.height=7, fig.align='centre'}
t_prob(2, 6, 'both')
```

What is the probability that the absolute value of T is greater than 2?

```{r t_prob4, fig.width=7, fig.height=7, fig.align='centre'}
t_prob(2, 6, 'interval')
```
