---
title: "Visualization"
author: "Aravind Hebbali"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE, message=FALSE}
library(descriptr)
library(dplyr)
```

## Data

We have modified the `mtcars` data to create a new data set `mtcarz`. The only
difference between the two data sets is related to the variable types. 

```{r egdata}
str(mtcarz)
```

### Continuous Data

The following functions will create plots for all or subset of continuous 
variables in the data set.

#### Histograms

```{r hist}
ds_plot_histogram(mtcarz)
```

#### Density Plots

```{r hist}
ds_plot_density(mtcarz)
```

#### Box Plots

```{r hist}
ds_plot_box_single(mtcarz)
```

#### Scatter Plots

```{r hist}
ds_plot_scatter(mtcarz, mpg, disp, hp)
```

### Categorical Data

The following functions will create plots for all or subset of categorical 
variables in the data set.

#### Bar Plot

```{r hist}
ds_plot_bar(mtcarz)
```

#### Stacked Bar Plot

```{r hist}
ds_plot_bar_stacked(mtcarz, cyl, gear, am)
```

#### Grouped Bar Plot

```{r hist}
ds_plot_bar_grouped(mtcarz, cyl, gear, am)
```

#### Grouped Box Plots

```{r hist}
ds_plot_box_group(mtcarz, cyl, gear, mpg, disp)
```
